\documentclass[10pt]{article}

\include{preamble}

\author{Sean B.S. Miller}
\date{\today}
\title{Notes on MIT lectures on\\ Effective Field Theory}

\begin{document}
	\maketitle
	\begin{abstract}
		abstract-text
	\end{abstract}
	\section{Lecture 2}
	Consider a scalar-field action:
	
	\begin{equation}
		S[\phi] = \int d^dx\left[ \frac{1}{2}\partial_\mu\phi\partial^\mu\phi - \frac{1}{2}m^2\phi^2 - \lambda\frac{1}{4!}\phi^4 - \tau\frac{1}{6!}\phi^6 + \ldots\right]
	\end{equation}
	
	where the values have mass dimensions:
	
	\begin{alignat}{4}
		& [\phi] &&= \frac{d-2}{2} \quad [d^dx] &&= -d \quad [m^2] &&= 2 \\
		& [\lambda] &&= 4-d \quad [\tau] &&= 6-2d \quad &&\ldots
	\end{alignat}
	
	Say we want to study the interaction matrix element $\langle\phi(x_1)\ldots\phi(x_n)\rangle$, at a long distance $x^\mu = Sx'^\mu$ where
	
	\begin{align}
		\begin{split}
		S&\rightarrow\infty \\
		x'&\rightarrow \text{fixed}
		\end{split}
	\end{align}
	
	Let
	
	\begin{equation}
		\phi(x) = S^{\frac{2-d}{2}}\phi'(x')
	\end{equation}
	
	\begin{equation}
		S'[\phi'] = \int d^dx'\left[ \frac{1}{2}\partial_\mu\phi'\partial^\mu\phi' - \frac{1}{2}m^2S^2\phi'^2 - \lambda\frac{S^{4-d}}{4!}\phi'^4 - \tau\frac{S^{6-2d}}{6!}\phi'^6 + \ldots\right]
	\end{equation}
	
	Now:
	
	\begin{equation}
		\langle\phi(S'x_1)\ldots\phi(Sx'_n)\rangle = S^{\frac{n(2-d)}{2}}\langle\phi'(x'_1)\ldots\phi'(x'_n)\rangle
	\end{equation}
	
	We know $x'$ is invariant with $S$, so the interaction elements in invariant under $S$. Assume $d=4$, when $S\rightarrow\infty$ we have the following properties:
	
	\begin{itemize}
		\item The $m^2$-term becomes increasingly important.
		\item The $\tau$-term becomes less important.
		\item The $\lambda$-term remains just as important.
	\end{itemize}
	
	Therefore:
	
	\begin{itemize}
		\item $\phi^2$ is relevant
		\item $\phi^4$ is marginal
		\item $\phi^6$ is irrelevant
	\end{itemize}
	
	For finite (large) $S$, the dimension of parameters (or operators) tells us their importance. Use the mass scale of parameters:
	
	\begin{align}
		m^2 &\sim \Lambda_{new}^{(2)} \\
		\lambda &\sim \Lambda_{new}^{(0)} \\
		\tau &\sim \Lambda_{new}^{(-2)}
	\end{align}
	
	Since large distances means small momenta, i.e. $p \ll \Lambda_{new}$.
	
	\subsection{Divergences}
	Take now $m=0$, or small such that $m^2S^2 \sim 1$.
	
	\begin{equation}
		\sim \lambda^2\int \dbar^dk\left(\frac{1}{k^2 - m^2 + i\epsilon}\right)\left(\frac{1}{(k+p)^2 - m^2 + i\epsilon}\right)
	\end{equation}
	
	which diverges as $\Lambda^{d-4}$, and $d-4$ is the degree of divergences.
	
	\begin{equation}
		d=4 \sim \int \frac{d^dk}{k^4} \sim \int \frac{dk}{k} \sim \ln(\Lambda) \leftrightarrow \epsilon^{-1}
	\end{equation}
	
	This renormalizes the $\lambda\phi^4$ interaction vertex.\\
	
	The loop diagrams with a $\lambda$ and $\tau$ vertex is similar to:
	
	\begin{equation}
		\lambda\tau\int \frac{\dbar^d k}{(\ldots)(\ldots)}	
	\end{equation}
	
	which renormalizes the $\tau\phi^6$-vertex.\\
	The loop diagrams with two $\tau$ vertices is similar to:
	
	\begin{equation}
	\tau^2\int \frac{\dbar^d k}{(\ldots)(\ldots)}	
	\end{equation}
	
	which renormalizes the $\phi^8$-vertex.\\
	Since $\phi^8$ is not in $S[\phi]$ (unless we consider the $(+\ldots)$ terms), the theory is not renormalized in the traditional sense. However, if $\tau\sim \Lambda_{new}^{(-2)}$ is small ($p^2\tau \ll 1$), then the theory can be renormalized order by order in $(\Lambda_{new}^{(-1)})$.\\
	To include all corrections up to $\Lambda_{new}^{(-r)}$ or $S^{-r}$, we include all operators with dimensions $[\mathcal{O}] \leq d+r$. Here we assume power counting is synonymous with dimensions (i.e. $x_i^\mu = Sx_i'^\mu$ where $S_i=S\:\forall\:i$, meaning $S$ is universal). For the standard model (SM), $\mathcal{L}^{(0)}$, all operators are of order $[\mathcal{O}]\leq 4$, and that it is renormalizable in the traditional sense. For SM correlations, $\mathcal{L}^{(1)}$, we add
	
	\begin{equation}
		\mathcal{L}^{(1)} = \frac{C}{\Lambda_{new}}\mathcal{O}_5 \quad,
	\end{equation}
	
	where $[\mathcal{O}_5] = 5$ and $[C] = 0$, meaning $C\sim 1$, and we have \emph{made} $\Lambda_{new}$ explicit.\\
	Since nothing in $\mathcal{L}^{(0)}$ constrains $\Lambda_{new}$, we're to take\footnote{Here, $m_t$ and $m_W$ are the top quark and $W$-boson masses, respectively.} $\Lambda_{new} \gg m_t,m_W$ by as much as we want.
	
	\subsection{Corrections to $\mathcal{L}^{(0)}$}
	
	\begin{equation}
	\mathcal{L}^{(0)} = \underset{\Lambda_{new}^{(0)}}{\mathcal{L}^{(0)}} + \underset{\Lambda_{new}^{(1)}}{\mathcal{L}^{(1)}} + \underset{\Lambda_{new}^{(2)}}{\mathcal{L}^{(2)}} + \ldots \quad \text{for}\quad p^2\sim m_t^2,\: m_t\ll\Lambda_{new}
	\end{equation}
	
	\begin{itemize}
		\item Assume the Lorentz and gauge invariances are unbroken. Each $\mathcal{L}^{(i)}$ is invariant under Lorentz and $SU(3)\times SU(2)\times U(1)$ transforms.
		\item Construct $\mathcal{L}^{(i)}$ from the same degrees of freedom as $\mathcal{L}^{(0)}$ and assume the Higgs vacuum expectation value ($v=246\:\text{GeV}$) holds.
		\item Assume no new particle produced at $p$, only at $\Lambda_{new}$.
	\end{itemize}
	
	Equation 1:
	
	\begin{equation}
		\mathcal{L}^{(1)} = \frac{C_5}{\Lambda_5}\epsilon_{ij}\bar{L}_L^{c_j}H^j\epsilon_{kl}L_L^kH^l
	\end{equation}
	
	where $\bar{L}_L^{c_j} \equiv \left(\bar{L}_L^{j}\right)^TC$, $H=\begin{pmatrix}h^+\\h^-\end{pmatrix}$, and $L_L=\begin{pmatrix}\nu_L\\e_L\end{pmatrix}$. This term is the \emph{only} dimension $-5$ operator consistent with symmetries.\\
	Replacing $H\rightarrow\begin{pmatrix}0\\\nu\end{pmatrix}$ gives Majorana mass terms for observed $\nu$:
	
	\begin{equation}
		\frac{1}{2}m_\nu\nu_L^a\nu_L^b\epsilon_{ab} + h.c. \quad\text{where}\quad m_\nu = \frac{C_5 v^2}{2\Lambda_{new}}
	\end{equation}
	
	Knowing that $m_\nu \leq 0.5\:\text{eV}$ leads us to believe $\Lambda_{new} \gtrsim 6\times 10^{14}\:\text{GeV}$ for $C_5\sim1$.\\
	
	Equation 2: Dimension $-6$ operators exists that violate baryon number.\\
	
	Equation 3: With lepton and baryon number imposed, there are 80 operators with dimension $-6$:
	
	\begin{equation}
		\mathcal{L}^{(2)} = \sum_{i=1}^{80}\frac{c_i}{\Lambda_{new}^2}\mathcal{O}_i^{(6)}
	\end{equation} 
	
	There are two points to remark:
	\begin{enumerate}
		\item For any observable, only a "few" terms contribute.
		\item For any new theory at $\Lambda_{new}$, a particular pattern of $c_i$'s are expected.
	\end{enumerate}
	
\end{document}