\documentclass[aps,prc,twocolumn,showpacs,floatfix,nofootinbib,preprintnumbers,superscriptaddress,amsmath,amssymb]{revtex4-1}
\newcommand{\OP}[1]{{\bf\widehat{#1}}}

\newcommand{\be}{\begin{equation}}

\newcommand{\ee}{\end{equation}}

\begin{document}

\pagestyle{plain}

\section*{Master Thesis project for Christian Fleischer: Quantum-mechanical studies of infinite matter, with an emphasis on the infinite electron gas in two and three dimensions}

\section*{Introduction to the the physics of the electron gas}
The electron gas is perhaps the only realistic model of a 
system of many interacting particles that allows for a solution
of the Hartree-Fock equations on a closed form. Furthermore, to first order in the interaction, one can also
compute on a closed form the total energy and several other properties of a many-particle systems. 
The model gives a very good approximation to the properties of valence electrons in metals.

The homogeneous electron gas is one of the few examples of a system of many
interacting particles that allows for a solution of the mean-field
Hartree-Fock equations on a closed form.  To first order in the
electron-electron interaction, this applies to ground state properties
like the energy and its pertinent equation of state as well.  The
homogeneus electron gas is a system of electrons that is not
influenced by external forces except by an attraction provided by a
uniform background of ions. These ions give rise to a uniform
background charge.  The ions are stationary and the system as a whole
is neutral.
Irrespective of this simplicity, this system, in both two and
three-dimensions, has eluded a proper description of correlations in
terms of various first principle methods, except perhaps for quantum
Monte Carlo methods. In particular, the diffusion Monte Carlo
calculations of "Ceperley":"http://journals.aps.org/prl/abstract/10.1103/PhysRevLett.45.566" 
and "Ceperley and Tanatar":"http://journals.aps.org/prb/abstract/10.1103/PhysRevB.39.5005" 
are presently still considered as the
best possible benchmarks for the two- and three-dimensional electron
gas. 



The electron gas, in 
two or three dimensions is thus interesting as a test-bed for 
electron-electron correlations. The three-dimensional 
electron gas is particularly important as a cornerstone 
of the local-density approximation in density-functional 
theory. In the physical world, systems 
similar to the three-dimensional electron gas can be 
found in, for example, alkali metals and doped 
semiconductors. Two-dimensional electron fluids are 
observed on metal and liquid-helium surfaces, as well as 
at metal-oxide-semiconductor interfaces. However, the Coulomb 
interaction has an infinite range, and therefore 
long-range correlations play an essential role in the
electron gas. 




At low densities, the electrons become 
localized and form a lattice. This so-called Wigner 
crystallization is a direct consequence 
of the long-ranged repulsive interaction. At higher
densities, the electron gas is better described as a
liquid.
When using, for example, Monte Carlo methods the electron gas must be approximated 
by a finite system. The long-range Coulomb interaction 
in the electron gas causes additional finite-size effects  that are not
present in other infinite systems like nuclear matter or neutron star matter.
This poses additional challenges to many-body methods when applied 
to the electron gas.



\section*{Many-body approaches to infinite matter}
For an infinite homogeneous system  like nuclear matter or the electron gas, 
the one-particle wave functions are given by plane wave functions
normalized to a volume $\Omega$ for a quadratic box with length
$L$\footnote{The limit $L\rightarrow \infty$ is to be taken after we
  have computed various expectation values. In our case we will
  however always deal with a fixed number of particles and finite size
  effects become important. }
\[
\psi_{{\bf k}m_s}({\bf r})= \frac{1}{\sqrt{\Omega}}\exp{(i{\bf kr})}\xi_{m_s}
\]
where ${\bf k}$ is the wave number and $\xi_{m_s}$ is a spin function
for either spin up or down
\[ 
\xi_{m_{s}=+1/2}=\left(\begin{array}{c} 1
  \\ 0 \end{array}\right) \hspace{0.5cm}
\xi_{m_{s}=-1/2}=\left(\begin{array}{c} 0 \\ 1 \end{array}\right).\]

The single-particle energies for the three-dimensional electron gas
are
\[    
\varepsilon_{n_{x}, n_{y}, n_{z}} = \frac{\hbar^{2}}{2m}\left( \frac{2\pi }{L}\right)^{2}(n_{x}^{2}+n_{y}^{2}+n_{z}^{2}),
\]
resulting in the magic numbers $2$, $14$, $38$, $54$, etc.  

In general terms, our Hamiltonian contains at most a two-body
interactions. In second quantization, we can write our Hamiltonian as
\begin{equation}
\hat{H}= \sum_{pq}\langle p | \hat{h}_0 | q \rangle a_p^{\dagger} a_q + \frac{1}{4}\sum_{pqrs}\langle pq |v| r s \rangle a_p^{\dagger} a_q^{\dagger} a_s a_r,
\label{eq:ourHamiltonian}
\end{equation} 
where the operator $\hat{h}_0$ denotes the single-particle
Hamiltonian, and the elements $\langle pq|v|rs\rangle$ are the
anti-symmetrized Coulomb interaction matrix elements.  Normal-ordering
with respect to a reference state $|\Phi_0\rangle$ yields
\begin{equation}
\hat{H}=E_0 + \sum_{pq}f_{pq}\lbrace a_p^{\dagger} a_q\rbrace + \frac{1}{4}\sum_{pqrs}\langle pq |v| r s \rangle \lbrace a_p^{\dagger} a_q^{\dagger} a_s a_r \rbrace,
\label{eq:normalorder}
\end{equation}
where $E_0=\langle\Phi_0| \hat{H}| \Phi_0\rangle$ is the reference energy
and we have introduced the so-called  Fock matrix element defined as
\begin{equation}
f_{pq} = \langle p|\hat{h}_0| q \rangle + \sum\limits_{i} \langle pi |v| qi\rangle.
\label{eq:fockelement}
\end{equation}
The curly brackets in Eq.~(\ref{eq:normalorder}) indicate that the
creation and annihilation operators are normal ordered.

The unperturbed part
of the Hamiltonian is defined as the sum over all the single-particle
operators $\hat{h}_0$, resulting in
\[
\hat{H}_{0}=\sum_i\langle i|\hat{h}_0|i \rangle= {\displaystyle\sum_{{\bf k}_{i}m_{s_i}}}
\frac{\hbar^{2}k_i^{2}}{2m}a_{{\bf k}_{i}m_{s_i}}^{\dagger}
a_{{\bf k}_{i}m_{s_i}}.
\]
We will throughout suppress, unless explicitely needed, all references
to the explicit quantum numbers ${\bf k}_{i}m_{s_i}$. The summation
index $i$ runs over all single-hole states up to the Fermi level.

The general anti-symmetrized two-body interaction matrix element 
\[
\langle pq |v| r s \rangle = \langle
\mathbf{k}_{p}m_{s_{p}}\mathbf{k}_{q}m_{s_{q}}|v|\mathbf{k}_{r}m_{s_{r}}\mathbf{k}_{s}m_{s_{s}}\rangle,
\]
is given by the following expression
  \begin{align}
    & \langle
    \mathbf{k}_{p}m_{s_{p}}\mathbf{k}_{q}m_{s_{q}}|v|\mathbf{k}_{r}m_{s_{r}}\mathbf{k}_{s}m_{s_{s}}\rangle
    \nonumber \\ =&
    \frac{e^{2}}{\Omega}\delta_{\mathbf{k}_{p}+\mathbf{k}_{q},
      \mathbf{k}_{r}+\mathbf{k}_{s}}\left\{
    \delta_{m_{s_{p}}m_{s_{r}}}\delta_{m_{s_{q}}m_{s_{s}}}(1-\delta_{\mathbf{k}_{p}\mathbf{k}_{r}})\frac{4\pi
    }{\mu^{2} + (\mathbf{k}_{r}-\mathbf{k}_{p})^{2}} \right. \nonumber
    \\ & \left. -
    \delta_{m_{s_{p}}m_{s_{s}}}\delta_{m_{s_{q}}m_{s_{r}}}(1-\delta_{\mathbf{k}_{p}\mathbf{k}_{s}})\frac{4\pi
    }{\mu^{2} + (\mathbf{k}_{s}-\mathbf{k}_{p})^{2}} \right\},
    \nonumber
  \end{align}
for the three-dimensional electron gas.  The energy per electron computed with
the reference Slater determinant can then be written as 
(using hereafter only atomic units, meaning that $\hbar = m = e = 1$)
\[
E_0/N=\frac{1}{2}\left[\frac{2.21}{r_s^2}-\frac{0.916}{r_s}\right],
\]
for the three-dimensional electron gas.  This will serve (with the
addition of a Yukawa term) as the first benchmark. 
The electron gas provides a very useful benchmark at the Hartree-Fock level since it provides
an analytical solution for the Hartree-Fock energy single-particle energy and the total  energy per particle. 
In addition to the above studies of the electron gas, it
is important to study properly the boundary conditions as well. Here
we will study so-called periodic boundary conditions following the
work of Refs.~\cite{gros1992,gros1996,lin2001}. See also the recent
work of Hagen {\em et al.} and Baardsen, see Refs.~\cite{hagen2014,baardsen2014}. 
One aim 
of this thesis is to implement so-called twisted boundary conditions,
first for the simple electron gas at the Hartree-Fock level. 
The next
step is to perform diffusion Monte Carlo calculations 
for electron gas in two and three dimensions  using the so-called full configuration interaction quantum Monte Carlo approach developed by Ali Alavi {\em et al} and studied in detail by a former master of science student of the computational physics group, Karl Leikanger, now PhD student at CTCC.
The studies can, if time allows, also be extended to studies of finite fermionic systems like electrons in oscillator traps, so-called quantum dots systems or atoms. 

The results will be published in scientific journals, if possible.

\section*{Progress plan and milestones}
The aims and progress plan of this thesis are as follows
\begin{itemize}
\item Spring 2016 and partly fall 2016:  The first step is to set up a program that uses cartesian coordinates in order to compute the Hartree-Fock 
energy of the electron gas with a Coulomb interaction with and without a screening term (Yukawa like potential). 
\item Fall 2016:  The next step is to include a proper treatment of periodic boundary conditions using the twisted boudary condition
approach described in Refs.~\cite{gros1992,gros1996,lin2001}.
\item Fall 2016: Thereafter, correlations at the electron gas level
  will be studied using a simpler Variational Monte Carlo approach.
\item Spring 2017: The final step is to write a code which employs the Full Configuration Interaction Quantum Monte Carlo method of Alavi {\em et al} for the electron gas in two and three dimensions and compare these results with other many-body methods like density functional theory, coupled cluster theory and many-body perturbation theory.
\item Spring 2017: The final step is to perform systematic studies of the infinite electron gas at various densities, both in two and three dimensions.
\item Spring 2017: The last part deals with a proper write-up of the thesis. 
\end{itemize}
 
The thesis is expected to be handed in May/June 2017.


\bibliographystyle{apsrev}
\begin{thebibliography}{83}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname bibnamefont\endcsname\relax
  \def\bibnamefont#1{#1}\fi
\expandafter\ifx\csname bibfnamefont\endcsname\relax
  \def\bibfnamefont#1{#1}\fi
\expandafter\ifx\csname citenamefont\endcsname\relax
  \def\citenamefont#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{\citenamefont{Burrows}(2013)}]{burrows2013}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Burrows}},
  \bibinfo{journal}{Rev. Mod. Phys.} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{245} (\bibinfo{year}{2013}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/RevModPhys.85.245}.

\bibitem[{\citenamefont{Weber}(1999)}]{weber1999}
\bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Weber}},
  \emph{\bibinfo{title}{Pulsars as Astrophysical Laboratories for Nuclear and
  Particle Physics}} (\bibinfo{publisher}{Institute of Physics Publishing,
  London}, \bibinfo{year}{1999}).

\bibitem[{\citenamefont{Heiselberg and Hjorth-Jensen}(2000)}]{hh2000}
\bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Heiselberg}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Hjorth-Jensen}},
  \bibinfo{journal}{Phys. Rep.} \textbf{\bibinfo{volume}{328}},
  \bibinfo{pages}{237} (\bibinfo{year}{2000}).

\bibitem[{\citenamefont{Lattimer and Prakash}(2007)}]{lattimer2007}
\bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Lattimer}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Prakash}},
  \bibinfo{journal}{Phys. Rep.} \textbf{\bibinfo{volume}{442}},
  \bibinfo{pages}{109} (\bibinfo{year}{2007}).

\bibitem[{\citenamefont{Sammarruca}(2010)}]{sammarruca2010}
\bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Sammarruca}},
  \bibinfo{journal}{International Journal of Modern Physics E}
  \textbf{\bibinfo{volume}{19}}, \bibinfo{pages}{1259} (\bibinfo{year}{2010}),
  \urlprefix\url{http://www.worldscientific.com/doi/abs/10.1142/S0218301310015874}.

\bibitem[{\citenamefont{Lattimer}(2012)}]{lattimer2012}
\bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Lattimer}},
  \bibinfo{journal}{Ann. Rev. Nucl. Part. Science}
  \textbf{\bibinfo{volume}{62}}, \bibinfo{pages}{485} (\bibinfo{year}{2012}).

\bibitem[{\citenamefont{Hebeler et~al.}(2013)\citenamefont{Hebeler, Lattimer,
  Pethick, and Schwenk}}]{hebeler2012f}
\bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Hebeler}},
  \bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Lattimer}},
  \bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Pethick}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Schwenk}},
  \bibinfo{journal}{Astrophys. J.} \textbf{\bibinfo{volume}{773}},
  \bibinfo{pages}{11} (\bibinfo{year}{2013}).

\bibitem[{\citenamefont{Alex~Brown}(2000)}]{brown2000}
\bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{Alex~Brown}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{5296} (\bibinfo{year}{2000}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.85.5296}.

\bibitem[{\citenamefont{Horowitz and Piekarewicz}(2001)}]{horowitz2001}
\bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Horowitz}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Piekarewicz}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{86}},
  \bibinfo{pages}{5647} (\bibinfo{year}{2001}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.86.5647}.

\bibitem[{\citenamefont{Gandolfi et~al.}(2012)\citenamefont{Gandolfi, Carlson,
  and Reddy}}]{gandolfi2012}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Carlson}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Reddy}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{032801} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.85.032801}.

\bibitem[{\citenamefont{Tsang et~al.}(2012)\citenamefont{Tsang, Stone, Camera,
  Danielewicz, Gandolfi, Hebeler, Horowitz, Lee, Lynch, Kohley
  et~al.}}]{tsang2012}
\bibinfo{author}{\bibfnamefont{M.~B.} \bibnamefont{Tsang}},
  \bibinfo{author}{\bibfnamefont{J.~R.} \bibnamefont{Stone}},
  \bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Camera}},
  \bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Danielewicz}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Hebeler}},
  \bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Horowitz}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Lee}},
  \bibinfo{author}{\bibfnamefont{W.~G.} \bibnamefont{Lynch}},
  \bibinfo{author}{\bibfnamefont{Z.}~\bibnamefont{Kohley}},
  \bibnamefont{et~al.}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{015803}
  (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.86.015803}.

\bibitem[{\citenamefont{Steiner and Gandolfi}(2012)}]{steiner2012}
\bibinfo{author}{\bibfnamefont{A.~W.} \bibnamefont{Steiner}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{108}},
  \bibinfo{pages}{081102} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.108.081102}.

\bibitem[{\citenamefont{Shlomo and Youngblood}(1993)}]{shlomo1993}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Shlomo}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{D.~H.} \bibnamefont{Youngblood}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{47}},
  \bibinfo{pages}{529} (\bibinfo{year}{1993}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.47.529}.

\bibitem[{\citenamefont{Abrahamyan et~al.}(2012)\citenamefont{Abrahamyan,
  Ahmed, Albataineh, Aniol, Armstrong, Armstrong, Averett, Babineau, Barbieri,
  Bellini et~al.}}]{abrahamyan2012}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Abrahamyan}},
  \bibinfo{author}{\bibfnamefont{Z.}~\bibnamefont{Ahmed}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Albataineh}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Aniol}},
  \bibinfo{author}{\bibfnamefont{D.~S.} \bibnamefont{Armstrong}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Armstrong}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Averett}},
  \bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{Babineau}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Barbieri}},
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Bellini}},
  \bibnamefont{et~al.} (\bibinfo{collaboration}{PREX Collaboration}),
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{108}},
  \bibinfo{pages}{112502} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.108.112502}.

\bibitem[{\citenamefont{Reinhard et~al.}(2013)\citenamefont{Reinhard,
  Piekarewicz, Nazarewicz, Agrawal, Paar, and Roca-Maza}}]{reinhard2013}
\bibinfo{author}{\bibfnamefont{P.-G.} \bibnamefont{Reinhard}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Piekarewicz}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Nazarewicz}},
  \bibinfo{author}{\bibfnamefont{B.~K.} \bibnamefont{Agrawal}},
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Paar}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{X.}~\bibnamefont{Roca-Maza}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{88}},
  \bibinfo{pages}{034325} (\bibinfo{year}{2013}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.88.034325}.

\bibitem[{\citenamefont{Erler et~al.}(2013)\citenamefont{Erler, Horowitz,
  Nazarewicz, Rafalski, and Reinhard}}]{erler2013}
\bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Erler}},
  \bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Horowitz}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Nazarewicz}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Rafalski}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{P.-G.} \bibnamefont{Reinhard}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{87}},
  \bibinfo{pages}{044320} (\bibinfo{year}{2013}).

\bibitem[{\citenamefont{Kortelainen et~al.}(2010)\citenamefont{Kortelainen,
  Lesinski, Mor\'e, Nazarewicz, Sarich, Schunck, Stoitsov, and
  Wild}}]{kortelainen2010}
\bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Kortelainen}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Lesinski}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Mor\'e}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Nazarewicz}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Sarich}},
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Schunck}},
  \bibinfo{author}{\bibfnamefont{M.~V.} \bibnamefont{Stoitsov}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Wild}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{82}},
  \bibinfo{pages}{024313} (\bibinfo{year}{2010}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.82.024313}.

\bibitem[{\citenamefont{Lunney et~al.}(2003)\citenamefont{Lunney, Pearson, and
  Thibault}}]{lunney2003}
\bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Lunney}},
  \bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Pearson}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Thibault}},
  \bibinfo{journal}{Rev. Mod. Phys.} \textbf{\bibinfo{volume}{75}},
  \bibinfo{pages}{1021} (\bibinfo{year}{2003}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/RevModPhys.75.1021}.

\bibitem[{\citenamefont{Day}(1967)}]{day1967}
\bibinfo{author}{\bibfnamefont{B.~D.} \bibnamefont{Day}},
  \bibinfo{journal}{Rev. Mod. Phys.} \textbf{\bibinfo{volume}{39}},
  \bibinfo{pages}{719} (\bibinfo{year}{1967}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/RevModPhys.39.719}.

\bibitem[{\citenamefont{Brueckner et~al.}(1954)\citenamefont{Brueckner,
  Levinson, and Mahmoud}}]{brueckner1954}
\bibinfo{author}{\bibfnamefont{K.~A.} \bibnamefont{Brueckner}},
  \bibinfo{author}{\bibfnamefont{C.~A.} \bibnamefont{Levinson}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{H.~M.}
  \bibnamefont{Mahmoud}}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{217} (\bibinfo{year}{1954}).

\bibitem[{\citenamefont{Brueckner}(1955)}]{brueckner1955}
\bibinfo{author}{\bibfnamefont{K.~A.} \bibnamefont{Brueckner}},
  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{100}},
  \bibinfo{pages}{36} (\bibinfo{year}{1955}).

\bibitem[{\citenamefont{Baldo and Burgio}(2012)}]{baldo2012}
\bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Baldo}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{G.~F.} \bibnamefont{Burgio}},
  \bibinfo{journal}{Reports on Progress in Physics}
  \textbf{\bibinfo{volume}{75}}, \bibinfo{pages}{026301}
  (\bibinfo{year}{2012}).

\bibitem[{\citenamefont{Baldo et~al.}(2012)\citenamefont{Baldo, Polls, Rios,
  Schulze, and Vida\~na}}]{baldo2012a}
\bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Baldo}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Polls}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Rios}},
  \bibinfo{author}{\bibfnamefont{H.-J.} \bibnamefont{Schulze}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{I.}~\bibnamefont{Vida\~na}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{86}},
  \bibinfo{pages}{064001} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.86.064001}.

\bibitem[{\citenamefont{K{\"u}mmel et~al.}(1978)\citenamefont{K{\"u}mmel,
  L{\"u}hrmann, and Zabolitzky}}]{kummel1978}
\bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{K{\"u}mmel}},
  \bibinfo{author}{\bibfnamefont{K.~H.} \bibnamefont{L{\"u}hrmann}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{J.~G.}
  \bibnamefont{Zabolitzky}}, \bibinfo{journal}{Phys. Rep.}
  \textbf{\bibinfo{volume}{36}}, \bibinfo{pages}{1} (\bibinfo{year}{1978}).

\bibitem[{\citenamefont{Day and Zabolitzky}(1981)}]{day1981}
\bibinfo{author}{\bibfnamefont{B.~D.} \bibnamefont{Day}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Zabolitzky}},
  \bibinfo{journal}{Nucl. Phys. A} \textbf{\bibinfo{volume}{366}},
  \bibinfo{pages}{221} (\bibinfo{year}{1981}).

\bibitem[{\citenamefont{Carlson et~al.}(2003)\citenamefont{Carlson, Morales,
  Pandharipande, and Ravenhall}}]{carlson2003}
\bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Carlson}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Morales}},
  \bibinfo{author}{\bibfnamefont{V.~R.} \bibnamefont{Pandharipande}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{D.~G.}
  \bibnamefont{Ravenhall}}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{68}}, \bibinfo{pages}{025802}
  (\bibinfo{year}{2003}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.68.025802}.

\bibitem[{\citenamefont{Gandolfi et~al.}(2009)\citenamefont{Gandolfi,
  Illarionov, Schmidt, Pederiva, and Fantoni}}]{gandolfi2009}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{A.~Y.} \bibnamefont{Illarionov}},
  \bibinfo{author}{\bibfnamefont{K.~E.} \bibnamefont{Schmidt}},
  \bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Pederiva}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Fantoni}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{79}},
  \bibinfo{pages}{054005} (\bibinfo{year}{2009}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.79.054005}.

\bibitem[{\citenamefont{Gezerlis and Carlson}(2010)}]{gezerlis2010}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Gezerlis}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Carlson}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{81}},
  \bibinfo{pages}{025803} (\bibinfo{year}{2010}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.81.025803}.

\bibitem[{\citenamefont{Lovato et~al.}(2012)\citenamefont{Lovato, Benhar,
  Fantoni, and Schmidt}}]{lovato2012}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Lovato}},
  \bibinfo{author}{\bibfnamefont{O.}~\bibnamefont{Benhar}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Fantoni}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{K.~E.} \bibnamefont{Schmidt}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{024003} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.85.024003}.

\bibitem[{\citenamefont{Gezerlis et~al.}(2013)\citenamefont{Gezerlis, Tews,
  Epelbaum, Gandolfi, Hebeler, Nogga, and Schwenk}}]{gezerlis2013}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Gezerlis}},
  \bibinfo{author}{\bibfnamefont{I.}~\bibnamefont{Tews}},
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Epelbaum}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Hebeler}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Nogga}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Schwenk}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{111}},
  \bibinfo{pages}{032501} (\bibinfo{year}{2013}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.111.032501}.

\bibitem[{\citenamefont{Dickhoff and Barbieri}(2004)}]{dickhoff2004}
\bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Dickhoff}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Barbieri}},
  \bibinfo{journal}{Progress in Particle and Nuclear Physics}
  \textbf{\bibinfo{volume}{52}}, \bibinfo{pages}{377 } (\bibinfo{year}{2004}),
  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0146641004000535}.

\bibitem[{\citenamefont{Som\`a and Bo\ifmmode~\dot{z}\else
  \.{z}\fi{}ek}(2008)}]{soma2008}
\bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Som\`a}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Bo\ifmmode~\dot{z}\else
  \.{z}\fi{}ek}}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{78}}, \bibinfo{pages}{054003}
  (\bibinfo{year}{2008}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.78.054003}.

\bibitem[{\citenamefont{Rios and Som\`a}(2012)}]{soma2012}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Rios}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Som\`a}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{108}},
  \bibinfo{pages}{012501} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.108.012501}.

\bibitem[{\citenamefont{Carbone et~al.}(2013)\citenamefont{Carbone, Polls, and
  Rios}}]{carbone2013}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Carbone}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Polls}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Rios}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{88}},
  \bibinfo{pages}{044302} (\bibinfo{year}{2013}).



\bibitem[{\citenamefont{Gros}(1992)}]{gros1992}
\bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Gros}},
  \bibinfo{journal}{Zeitschrift f{\"u}r Physik B Condensed Matter}
  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{359} (\bibinfo{year}{1992}),
  ISSN \bibinfo{issn}{0722-3277},
  \urlprefix\url{http://dx.doi.org/10.1007/BF01323728}.

\bibitem[{\citenamefont{Gros}(1996)}]{gros1996}
\bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Gros}}, \bibinfo{journal}{Phys.
  Rev. B} \textbf{\bibinfo{volume}{53}}, \bibinfo{pages}{6865}
  (\bibinfo{year}{1996}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevB.53.6865}.

\bibitem[{\citenamefont{Lin et~al.}(2001)\citenamefont{Lin, Zong, and
  Ceperley}}]{lin2001}
\bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Lin}},
  \bibinfo{author}{\bibfnamefont{F.~H.} \bibnamefont{Zong}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{D.~M.} \bibnamefont{Ceperley}},
  \bibinfo{journal}{Phys. Rev. E} \textbf{\bibinfo{volume}{64}},
  \bibinfo{pages}{016702} (\bibinfo{year}{2001}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevE.64.016702}.

\bibitem{hagen2014} G.~Hagen {\em et al}, Phys.~Rev.~C {\bf 89}, 01431a9 (2014).

\bibitem{baardsen2014} G.~Baardsen, PhD thesis, University of Oslo (2014) 
\url{https://www.duo.uio.no/handle/10852/41025}. 

\bibitem[{\citenamefont{Shavitt and Bartlett}(2009)}]{shavittbartlett2009}
\bibinfo{author}{\bibfnamefont{I.}~\bibnamefont{Shavitt}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{R.~J.} \bibnamefont{Bartlett}},
  \emph{\bibinfo{title}{Many-body Methods in Chemistry and Physics}}
  (\bibinfo{publisher}{Cambridge University Press}, \bibinfo{year}{2009}).
\end{thebibliography}



\end{document}















