\documentclass[../main.tex]{subfiles}
 
\begin{document}

\begin{thebibliography}{99}

\bibitem{Manninen} S. M. Reimann and M. Manninen. Electronic structure of quantum dots. \emph{Reviews of Modern Physics}, \textbf{74}:1238, 2002. DOI: \url{https://doi.org/10.1103/RevModPhys.74.1283}.

\bibitem{QDotSolar} Z. Zheng, H. Ji, P. Yu and Z. Wang. Recent Progress Towards Quantum Dot Solar Cells with Enhanced Optical Absorption. \emph{Nanoscale Research Letters}, \textbf{11}:266, 2016. DOI: \url{https://doi.org/10.1186/s11671-016-1457-y}.

\bibitem{QDotSolar2} X. Lan, O. Voznyy, F. Pelayo García de Arquer, M. Liu, J. Xu, A. H. Proppe, G. Walters, F. Fan, H. Tan, M. Liu, Z. Yang, S. Hoogland, and E. H. Sargent. 10.6\% Certified Colloidal Quantum Dot Solar Cells via Solvent-Polarity-Engineered Halide Passivation. \emph{Nano Letters} \textbf{16}(7):4630-4634, 2016. DOI: \url{https://doi.org/10.1021/acs.nanolett.6b01957}.

\bibitem{QDotMed} S. Jin, Y. Hu, Z. Gu, L. Liu, and H. Wu. Application of Quantum Dots in Biological Imaging. \emph{Journal of Nanomaterials} \textbf{2011}:834139, 2011. DOI: \url{https://doi.org/10.1155/2011/834139}.

\bibitem{QDotMed2} L. A. Bentolila, X. Michalet, F. F. Pinaud, J. M. Tsay, S. Doose, J. J. Li, G. Sundaresan, A. M. Wu, S. S. Gambhir, S. Weiss. Quantum dots for molecular imaging and cancer medicine. \emph{Discovery Medicine} \textbf{5}(26):213–218, 2005. \url{https://www.ncbi.nlm.nih.gov/pubmed/20704913}

\bibitem{QDotLaser} N. N. Ledentsov. Quantum dot laser. \emph{Semiconductor Science and Technology} \textbf{26}:014001, 2011. DOI: \url{https://doi.org/10.1088/0268-1242/26/1/014001}.

\bibitem{QDotLaser2} V. H. Iyer, R. Mahadevu, and A. Pandey. Low Threshold Quantum Dot Lasers. \emph{The Journal of Physical Chemistry Letters} \textbf{7}(7):1244-1248, 2016. DOI: \url{https://doi.org/10.1021/acs.jpclett.6b00430}.

\bibitem{QDotComp} H. Wei and F. Deng. Scalable quantum computing based on stationary spin qubits in coupled quantum dots inside double-sided optical microcavities. \emph{Scientific Reports} \textbf{4}:7551, 2014. DOI: \url{https://doi.org/10.1038/srep07551}.

\bibitem{QDotComp2} M. A. Eriksson, M. Friesen, S. N. Coppersmith, R. Joynt, L. J. Klein, K. Slinker, C. Tahan, P. M. Mooney, J. O. Chu and S. J. Koester. Spin-Based Quantum Dot Quantum Computing in Silicon.\emph{Quantum Information Processing} \textbf{3}:133, 2004. DOI: \url{https://doi.org/10.1007/s11128-004-2224-z}.

\bibitem{QMC1} F. Bolton. Fixed-phase quantum Monte Carlo method applied to interacting electrons in
a quantum dot. \emph{Physical Review B} \textbf{54}:4780, 1996. DOI: \url{https://doi.org/10.1103/PhysRevB.54.4780}.

\bibitem{QMC2} F. Pederiva, C. J. Umrigar, and E. Lipparini. Diffusion Monte Carlo study of
circular quantum dots. \emph{Physical Review B} \textbf{62}:8120, 2000. DOI: \url{https://doi.org/10.1103/PhysRevB.62.8120}.

\bibitem{QMC3} A. D. Güçlü, Jian-Sheng Wang, and Hong Guo. Disordered quantum dots: A diffusion
quantum Monte Carlo study. \emph{Physical Review B} \textbf{68}:035304, 2003. DOI: \url{https://doi.org/10.1103/PhysRevB.68.035304}.

\bibitem{QDotBenchmarks} M.~.L.~Pedersen, G.~Hagen, M.~Hjorth-Jensen, S.~Kvaal, and F.~Pederiva. Ab initio computation of the energies of circular quantum dots. \emph{Physical Review B} {\bf 84}:115302, 2011. DOI: \url{https://doi.org/10.1103/PhysRevB.84.115302}.

\bibitem{ultra-cold neutrons} P. R. Huffman, C. R. Brome, J. S. Butterworth, K. J. Coakley, M. S. Dewey, S. N. Dzhosyuk, R. Golub, G. L. Greene, K. Habicht, S. K. Lamoreaux, C. E. H. Mattoni, D. N. McKinsey, F. E. Wietfeldt and J. M. Doyle. Magnetic trapping of neutrons. \emph{Nature} \textbf{403}:62-64, 2000. DOI: \url{https://doi.org/10.1038/47444}.
%Neutron Trapping Demonstrated for First Time at NIST. January 05, 2000. \url{https://www.nist.gov/news-events/news/2000/01/neutron-trapping-demonstrated-first-time-nist}, Read 29.05.2017.

\bibitem{Gardiner} C. W. Gardiner. \emph{Handbook of Stochastic Methods for Physics, Chemistry and the Natural Sciences}, Springer, 2nd edition, 1985.

\bibitem{Flyvbjerg} H. Flyvbjerg and H. G. Petersen. Error estimates on averages of correlated data. \emph{The Journal of Chemical Physics} \textbf{91}:461, 1989. DOI: \url{https://doi.org/10.1063/1.457480}.

%\bibitem{FYS4411-CG} M.~Hjorth-Jensen, {\em Conjugate gradient methods and other optimization methods},  \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/master/doc/pub/cg/pdf/cg-print.pdf}, Read: 17.03.2016.

\bibitem{Kalos} J. W. Moskowitz, M. H. Kalos. A new look at correlations in atomic and molecular systems. I. Application of fermion monte carlo variational method. \emph{International Journal of Quantum Chemistry} \textbf{20}:1107, 1981. DOI: \url{https://doi.org/10.1002/qua.560200508}.

\bibitem{Taut} M.~Taut. Two electrons in an external oscillator potential: Particular analytic solutions
of a Coulomb correlation problem. \emph{Physical Review A} {\bf 48}:3561-3566, 1993. DOI: \url{https://doi.org/10.1103/PhysRevA.48.3561}.

%\bibitem{FYS4411-Slides} M.~Hjorth-Jensen, \emph{Computational Physics 2: Variational Monte Carlo methods}, \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/master/doc/pub/vmc/pdf/vmc-print.pdf}, Read: 16.06.2016.

\bibitem{FYS4411-LectureNotes} M.~Hjorth-Jensen, \emph{Computational Physics Lecture Notes Fall 2015},  \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/master/doc/Literature/lectures2015.pdf}.%, Chapter 16.10, Read: 16.06.2016.

\bibitem{Griffiths} D. J. Griffiths, \emph{Introduction to Quantum Mechanics}, Pearson,
2nd edition, 2005.

\bibitem{Scientific Computing} A. Tveito, H. P. Langtangen, B. F. Nielsen and X. Cai, \emph{Elements of Scientific Computing}, Springer, 1st edition, 2010.

\bibitem{Hermite} B.H. Bransden and C.J. Joachain, \emph{Physics of Atoms and molecules}, Longman, 1983. Chapter
2.

\bibitem{Schrodinger} Georgia State University, \emph{Time Independent Schrodinger Equation}, Hyperphysics entry, \url{http://hyperphysics.phy-astr.gsu.edu/hbase/quantum/Scheq.html}, Read 23.05.17.

%\bibitem{CP1_Project 2} M.~Hjorth-Jensen, \url{https://github.com/CompPhysics/ComputationalPhysics/blob/master/doc/Projects/2015/Project2/project2_2015.pdf}

\bibitem{master project} M.~Hjorth-Jensen. \url{https://github.com/CompPhysics/ThesisProjects/blob/master/doc/pub/quantumdots/pdf/quantumdots-minted.pdf}.

%\bibitem{basicMB} M.~Hjorth-Jensen, \emph{Hartree-Fock methods and introduction to Many-body Theory}, \url{https://github.com/CompPhysics/ComputationalPhysics2/blob/gh-pages/doc/pub/basicMB/pdf/basicMB-print.pdf}

\bibitem{Ruder} S. Ruder, An overview of gradient descent optimization
algorithms. \href{https://arxiv.org/abs/1609.04747}{arXiv:1609.04747 [cs.LG]}, 2016.

%\bibitem{QD-wiki} \url{https://en.wikipedia.org/wiki/Quantum_dot}, Read: 16.06.2016.

\bibitem{Stroustrup} B. Stroustrup, \emph{Bjarne Stroustrup's C++ Glossary}.  \url{http://www.stroustrup.com/glossary.html#Gpolymorphism}, Read: 09.05.2017.

\bibitem{gccDocs} \emph{Using the GNU Compiler Collection (GCC)}. section 3.10: Options That Control Optimization. \url{http://gcc.gnu.org/onlinedocs/gcc/Optimize-Options.html}, Read 10.05.2017.

\bibitem{github} Christian Fleishcer, Github Repository 2017,. \url{https://github.com/christianfleischer/Quantum-Dot-Project}.

\bibitem{Blaise} B. Barney. Lawrence Livermore National Laboratory. \emph{Introduction to Parallel Computing}. \url{https://computing.llnl.gov/tutorials/parallel_comp/}, Read 26.05.2017.

\bibitem{Wheeler} D. A. Wheeler. \emph{Secure Programming HOWTO}, Chapter 7.  \url{https://www.dwheeler.com/secure-programs/Secure-Programs-HOWTO/avoid-race.html}, Read 23.05.2017.

\bibitem{Lin} C. Lin. \emph{Logic Gates}. computer organization class notes, 2003. \url{https://www.cs.umd.edu/class/sum2003/cmsc311/Notes/Comb/gates.html}, Read 23.05.2017.

\bibitem{Yang} Y. M. Wang. Coupled-Cluster Studies of Double Quantum
Dots. Master’s thesis, University of Oslo, 2011. \url{http://urn.nb.no/URN:NBN:no-28527}.

\bibitem{Jorgen} J. Høgberget. Quantum Monte-Carlo Studies of Generalized Many-body Systems. 
Master’s thesis, University of Oslo, 2013. \url{http://urn.nb.no/URN:NBN:no-38645}.

\bibitem{Reimann} S. Reimann. Quantum-mechanical systems in traps and Similarity Renormalization Group theory. Master’s thesis, University of Oslo, 2013. \url{http://urn.nb.no/URN:NBN:no-38639}.

\bibitem{Hirth} C. Hirth. Studies of quantum dots: Ab initio coupledcluster analysis using OpenCL and GPU programming. Master’s thesis, University of Oslo, 2012. \url{http://urn.nb.no/URN:NBN:no-31657}.

\bibitem{Olsen} V. K. B. Olsen. Full Configuration Interaction Simulation of Quantum Dots. Master’s thesis, University of Oslo, 2012. \url{http://urn.nb.no/URN:NBN:no-32952}.

\bibitem{Armadillo} C. Sanderson and R. Curtin. Armadillo: a template-based C++ library for linear algebra. 
\emph{Journal of Open Source Software} \textbf{1}:26, 2016. \url{http://dx.doi.org/10.21105/joss.00026}.

\bibitem{Open MPI} E. Gabriel, G. E. Fagg, G. Bosilca, T. Angskun, J. J. Dongarra, J. M. Squyres, V. Sahay, P. Kambadur, B. Barrett, A. Lumsdaine, R. H. Castain, D. J. Daniel, R. L. Graham, and T. S. Woodall. Open MPI: Goals, Concept, and Design of a Next Generation MPI Implementation. In Proceedings, 11th European PVM/MPI Users' Group Meeting, Budapest, Hungary, September 2004.

\bibitem{Matplotlib} J. D. Hunter. Matplotlib: A 2D Graphics Environment. \emph{Computing in Science \& Engineering} \textbf{9}:90-95, 2007. DOI: \url{http://dx.doi.org/10.1109/MCSE.2007.55}, Publisher link: \url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2007.55}.

\bibitem{NumPy} S. van der Walt, S. C. Colbert and G. Varoquaux. The NumPy Array: A Structure for Efficient Numerical Computation. \emph{Computing in Science \& Engineering} \textbf{13}:22-30, 2011.  DOI: \url{http://dx.doi.org/10.1109/MCSE.2011.37}, Publisher link: \url{http://aip.scitation.org/doi/abs/10.1109/MCSE.2011.37}.

\bibitem{SymPy} A. Meurer, C. P. Smith, M. Paprocki, O. Čertík, S. B. Kirpichev, M. Rocklin, A. Kumar, S. Ivanov, J. K. Moore, S. Singh, T. Rathnayake, S. Vig, B. E. Granger, R. P. Muller, F. Bonazzi, H. Gupta, S. Vats, F. Johansson, F. Pedregosa, M. J. Curry, A. R. Terrel, Š. Roučka, A. Saboo, I. Fernando, S. Kulal, R. Cimrman, A. Scopatz. SymPy: symbolic computing in Python. \emph{PeerJ Computer Science} \textbf{3}:e103, 2017. DOI: \url{https://doi.org/10.7717/peerj-cs.103}.

\bibitem{Multiwells} D. V. Schroeder, \emph{Multiple Wells} \url{http://physics.weber.edu/schroeder/quantum/MultipleWells.pdf}, Read 26.05.2017.



\end{thebibliography}
\end{document}