\documentclass[aps,prc,twocolumn,showpacs,floatfix,nofootinbib,preprintnumbers,superscriptaddress,amsmath,amssymb]{revtex4-1}
\newcommand{\OP}[1]{{\bf\widehat{#1}}}

\newcommand{\be}{\begin{equation}}

\newcommand{\ee}{\end{equation}}

\begin{document}

\pagestyle{plain}

\section*{Master Thesis project for Hans Mathias Mamen Vege: Quantum-mechanical studies of infinite nuclear matter and neutron star matter and neutrinos}


\section{Introduction}

Bulk nucleonic matter is interesting for several reasons. The equation of state (EoS) of
neutron matter determines properties of supernova
explosions~\cite{burrows2013}, and of neutron
stars~\cite{weber1999,hh2000,lattimer2007,sammarruca2010,lattimer2012,hebeler2012f},
and it links the latter to neutron radii in atomic
nuclei~\cite{brown2000,horowitz2001,gandolfi2012} and the symmetry
energy~\cite{tsang2012,steiner2012}. Similarly, the compressibility of
nuclear matter is probed in giant dipole
excitations~\cite{shlomo1993}, and the symmetry energy of nuclear
matter is related to the difference between proton and neutron radii
in atomic nuclei~\cite{abrahamyan2012,reinhard2013,erler2013}. The
saturation point of nuclear matter determines bulk properties of
atomic nuclei, and is therefore an important constraint for nuclear
energy-density functionals and mass models (see,
e.g., Refs.~\cite{kortelainen2010,lunney2003}).

The determination and our understanding of the EoS for nuclear matter
is intimately linked with our capability to solve the nuclear
many-body problem. Here, correlations beyond the mean field play an
important role.  Theoretical studies of nuclear matter and the
pertinent EoS span back to the very early days of nuclear many-body
physics. Early computations are nicely described in the 1967 review by
Day \cite{day1967}. These early calculations were performed using
Brueckner-Bethe-Goldstone theory \cite{brueckner1954,brueckner1955},
see Refs.\cite{hh2000,baldo2012,baldo2012a,baardsen2014} for recent reviews and
developments.  In these calculations, mainly particle-particle
correlations were summed to infinite order.  Other correlations were
often included in a perturbative way. Coupled-cluster calculations of
nuclear matter were performed already during the late 1970s and early
1980s~\cite{kummel1978,day1981}. In recent years, there has been a
considerable algorithmic development of first-principle methods for
solving the nuclear many-body problem. A systematic inclusion of other
correlations in a non-perturbative way are nowadays accounted for in
Monte Carlo methods
\cite{carlson2003,gandolfi2009,gezerlis2010,lovato2012,gezerlis2013},
self-consistent Green's function approaches
\cite{dickhoff2004,soma2008, soma2012, baldo2012a,carbone2013} and
nuclear density functional theory \cite{lunney2003,erler2013}.


\section*{Many-body approaches to infinite matter}

This project is based on existing many-body codes for infinite matter using coupled cluster theory, see
for example the recent article in \url{http://lanl.arxiv.org/abs/1611.06765}. 

The first step of the thesis is to perform a coupled-cluster calculations at the
coupled-cluster with doubles excitations, the so-called CCD approach \cite{shavittbartlett2009}, using the nuclear matter code from \url{http://lanl.arxiv.org/abs/1611.06765}.  The next step
is to include triples correlations and perform full doubles and
triples calculations (CCDT) for neutron matter using a simplified model for the nuclear force, the so-called Minnesota potential. 
This allows for a benchmark of codes to infinite nuclear matter. After having developed the necessary formalism, the next step 
 is to include realistic models for the nuclear forces and 
study the EoS for pure neutron matter, asymmetric
nuclear matter (for different proton fractions) and symmetric nuclear
matter. With this one can study $\beta$-stable neutron star matter and
extract important information about the symmetry energy in infinite
matter and the composition of a neutron star. 
The resulting effective interactions at the two-body level can then be included in the study of neutrino emissivities in dense matter represented by the neutrino production processes 
\begin{equation}
    n+n\rightarrow p+n +e +\overline{\nu}_e,
    \hspace{0.5cm} p+n+e \rightarrow
    n+n+\nu_e .
    \label{eq:ind_neutr}
\end{equation}
These reactions correspond to the processes for 
$\beta$-decay and electron capture with a bystanding neutron.
The calculation of neutrino spectra has important consequences for our basic understanding on how neutron stars cool, the synthesis of the elements and neutrino oscillations in dense matter and an precise determination of neutrino spectra is the final goal of this thesis project.


The results will be published in scientific journals, if possible.

\section*{Progress plan and milestones}
The aims and progress plan of this thesis are as follows
\begin{itemize}
\item Spring 2017 and partly fall 2017:  Finalize courses and start developing a coupled cluster for infinite nuclear matter. 
\item Fall 2017:  The next step is to include a proper treatment of periodic boundary conditions using the twisted boudary condition
approach described in Refs.~\cite{gros1992,gros1996,lin2001}, triples correlations and realistic nuclear forces.
\item Spring 2018: With a functioning code the final aim is to perform systematic studies of  $\beta$-stable nuclear matter and to estimate the production rates of neutrinos in dense matter by studying various neutrino emitting processes. 
\item Spring 2018: The last part deals with a proper write-up of the thesis. 
\end{itemize}
 
The thesis is expected to be handed in May/June 2018.


\bibliographystyle{apsrev}
\begin{thebibliography}{83}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\expandafter\ifx\csname bibnamefont\endcsname\relax
  \def\bibnamefont#1{#1}\fi
\expandafter\ifx\csname bibfnamefont\endcsname\relax
  \def\bibfnamefont#1{#1}\fi
\expandafter\ifx\csname citenamefont\endcsname\relax
  \def\citenamefont#1{#1}\fi
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{\citenamefont{Burrows}(2013)}]{burrows2013}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Burrows}},
  \bibinfo{journal}{Rev. Mod. Phys.} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{245} (\bibinfo{year}{2013}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/RevModPhys.85.245}.

\bibitem[{\citenamefont{Weber}(1999)}]{weber1999}
\bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Weber}},
  \emph{\bibinfo{title}{Pulsars as Astrophysical Laboratories for Nuclear and
  Particle Physics}} (\bibinfo{publisher}{Institute of Physics Publishing,
  London}, \bibinfo{year}{1999}).

\bibitem[{\citenamefont{Heiselberg and Hjorth-Jensen}(2000)}]{hh2000}
\bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Heiselberg}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Hjorth-Jensen}},
  \bibinfo{journal}{Phys. Rep.} \textbf{\bibinfo{volume}{328}},
  \bibinfo{pages}{237} (\bibinfo{year}{2000}).

\bibitem[{\citenamefont{Lattimer and Prakash}(2007)}]{lattimer2007}
\bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Lattimer}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Prakash}},
  \bibinfo{journal}{Phys. Rep.} \textbf{\bibinfo{volume}{442}},
  \bibinfo{pages}{109} (\bibinfo{year}{2007}).

\bibitem[{\citenamefont{Sammarruca}(2010)}]{sammarruca2010}
\bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Sammarruca}},
  \bibinfo{journal}{International Journal of Modern Physics E}
  \textbf{\bibinfo{volume}{19}}, \bibinfo{pages}{1259} (\bibinfo{year}{2010}),
  \urlprefix\url{http://www.worldscientific.com/doi/abs/10.1142/S0218301310015874}.

\bibitem[{\citenamefont{Lattimer}(2012)}]{lattimer2012}
\bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Lattimer}},
  \bibinfo{journal}{Ann. Rev. Nucl. Part. Science}
  \textbf{\bibinfo{volume}{62}}, \bibinfo{pages}{485} (\bibinfo{year}{2012}).

\bibitem[{\citenamefont{Hebeler et~al.}(2013)\citenamefont{Hebeler, Lattimer,
  Pethick, and Schwenk}}]{hebeler2012f}
\bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Hebeler}},
  \bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Lattimer}},
  \bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Pethick}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Schwenk}},
  \bibinfo{journal}{Astrophys. J.} \textbf{\bibinfo{volume}{773}},
  \bibinfo{pages}{11} (\bibinfo{year}{2013}).

\bibitem[{\citenamefont{Alex~Brown}(2000)}]{brown2000}
\bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{Alex~Brown}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{5296} (\bibinfo{year}{2000}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.85.5296}.

\bibitem[{\citenamefont{Horowitz and Piekarewicz}(2001)}]{horowitz2001}
\bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Horowitz}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Piekarewicz}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{86}},
  \bibinfo{pages}{5647} (\bibinfo{year}{2001}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.86.5647}.

\bibitem[{\citenamefont{Gandolfi et~al.}(2012)\citenamefont{Gandolfi, Carlson,
  and Reddy}}]{gandolfi2012}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Carlson}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Reddy}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{032801} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.85.032801}.

\bibitem[{\citenamefont{Tsang et~al.}(2012)\citenamefont{Tsang, Stone, Camera,
  Danielewicz, Gandolfi, Hebeler, Horowitz, Lee, Lynch, Kohley
  et~al.}}]{tsang2012}
\bibinfo{author}{\bibfnamefont{M.~B.} \bibnamefont{Tsang}},
  \bibinfo{author}{\bibfnamefont{J.~R.} \bibnamefont{Stone}},
  \bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Camera}},
  \bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Danielewicz}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Hebeler}},
  \bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Horowitz}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Lee}},
  \bibinfo{author}{\bibfnamefont{W.~G.} \bibnamefont{Lynch}},
  \bibinfo{author}{\bibfnamefont{Z.}~\bibnamefont{Kohley}},
  \bibnamefont{et~al.}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{015803}
  (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.86.015803}.

\bibitem[{\citenamefont{Steiner and Gandolfi}(2012)}]{steiner2012}
\bibinfo{author}{\bibfnamefont{A.~W.} \bibnamefont{Steiner}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{108}},
  \bibinfo{pages}{081102} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.108.081102}.

\bibitem[{\citenamefont{Shlomo and Youngblood}(1993)}]{shlomo1993}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Shlomo}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{D.~H.} \bibnamefont{Youngblood}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{47}},
  \bibinfo{pages}{529} (\bibinfo{year}{1993}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.47.529}.

\bibitem[{\citenamefont{Abrahamyan et~al.}(2012)\citenamefont{Abrahamyan,
  Ahmed, Albataineh, Aniol, Armstrong, Armstrong, Averett, Babineau, Barbieri,
  Bellini et~al.}}]{abrahamyan2012}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Abrahamyan}},
  \bibinfo{author}{\bibfnamefont{Z.}~\bibnamefont{Ahmed}},
  \bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{Albataineh}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Aniol}},
  \bibinfo{author}{\bibfnamefont{D.~S.} \bibnamefont{Armstrong}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Armstrong}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Averett}},
  \bibinfo{author}{\bibfnamefont{B.}~\bibnamefont{Babineau}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Barbieri}},
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Bellini}},
  \bibnamefont{et~al.} (\bibinfo{collaboration}{PREX Collaboration}),
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{108}},
  \bibinfo{pages}{112502} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.108.112502}.

\bibitem[{\citenamefont{Reinhard et~al.}(2013)\citenamefont{Reinhard,
  Piekarewicz, Nazarewicz, Agrawal, Paar, and Roca-Maza}}]{reinhard2013}
\bibinfo{author}{\bibfnamefont{P.-G.} \bibnamefont{Reinhard}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Piekarewicz}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Nazarewicz}},
  \bibinfo{author}{\bibfnamefont{B.~K.} \bibnamefont{Agrawal}},
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Paar}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{X.}~\bibnamefont{Roca-Maza}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{88}},
  \bibinfo{pages}{034325} (\bibinfo{year}{2013}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.88.034325}.

\bibitem[{\citenamefont{Erler et~al.}(2013)\citenamefont{Erler, Horowitz,
  Nazarewicz, Rafalski, and Reinhard}}]{erler2013}
\bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Erler}},
  \bibinfo{author}{\bibfnamefont{C.~J.} \bibnamefont{Horowitz}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Nazarewicz}},
  \bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Rafalski}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{P.-G.} \bibnamefont{Reinhard}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{87}},
  \bibinfo{pages}{044320} (\bibinfo{year}{2013}).

\bibitem[{\citenamefont{Kortelainen et~al.}(2010)\citenamefont{Kortelainen,
  Lesinski, Mor\'e, Nazarewicz, Sarich, Schunck, Stoitsov, and
  Wild}}]{kortelainen2010}
\bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Kortelainen}},
  \bibinfo{author}{\bibfnamefont{T.}~\bibnamefont{Lesinski}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Mor\'e}},
  \bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Nazarewicz}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Sarich}},
  \bibinfo{author}{\bibfnamefont{N.}~\bibnamefont{Schunck}},
  \bibinfo{author}{\bibfnamefont{M.~V.} \bibnamefont{Stoitsov}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Wild}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{82}},
  \bibinfo{pages}{024313} (\bibinfo{year}{2010}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.82.024313}.

\bibitem[{\citenamefont{Lunney et~al.}(2003)\citenamefont{Lunney, Pearson, and
  Thibault}}]{lunney2003}
\bibinfo{author}{\bibfnamefont{D.}~\bibnamefont{Lunney}},
  \bibinfo{author}{\bibfnamefont{J.~M.} \bibnamefont{Pearson}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Thibault}},
  \bibinfo{journal}{Rev. Mod. Phys.} \textbf{\bibinfo{volume}{75}},
  \bibinfo{pages}{1021} (\bibinfo{year}{2003}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/RevModPhys.75.1021}.

\bibitem[{\citenamefont{Day}(1967)}]{day1967}
\bibinfo{author}{\bibfnamefont{B.~D.} \bibnamefont{Day}},
  \bibinfo{journal}{Rev. Mod. Phys.} \textbf{\bibinfo{volume}{39}},
  \bibinfo{pages}{719} (\bibinfo{year}{1967}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/RevModPhys.39.719}.

\bibitem[{\citenamefont{Brueckner et~al.}(1954)\citenamefont{Brueckner,
  Levinson, and Mahmoud}}]{brueckner1954}
\bibinfo{author}{\bibfnamefont{K.~A.} \bibnamefont{Brueckner}},
  \bibinfo{author}{\bibfnamefont{C.~A.} \bibnamefont{Levinson}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{H.~M.}
  \bibnamefont{Mahmoud}}, \bibinfo{journal}{Phys. Rev.}
  \textbf{\bibinfo{volume}{95}}, \bibinfo{pages}{217} (\bibinfo{year}{1954}).

\bibitem[{\citenamefont{Brueckner}(1955)}]{brueckner1955}
\bibinfo{author}{\bibfnamefont{K.~A.} \bibnamefont{Brueckner}},
  \bibinfo{journal}{Phys. Rev.} \textbf{\bibinfo{volume}{100}},
  \bibinfo{pages}{36} (\bibinfo{year}{1955}).

\bibitem[{\citenamefont{Baldo and Burgio}(2012)}]{baldo2012}
\bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Baldo}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{G.~F.} \bibnamefont{Burgio}},
  \bibinfo{journal}{Reports on Progress in Physics}
  \textbf{\bibinfo{volume}{75}}, \bibinfo{pages}{026301}
  (\bibinfo{year}{2012}).

\bibitem[{\citenamefont{Baldo et~al.}(2012)\citenamefont{Baldo, Polls, Rios,
  Schulze, and Vida\~na}}]{baldo2012a}
\bibinfo{author}{\bibfnamefont{M.}~\bibnamefont{Baldo}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Polls}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Rios}},
  \bibinfo{author}{\bibfnamefont{H.-J.} \bibnamefont{Schulze}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{I.}~\bibnamefont{Vida\~na}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{86}},
  \bibinfo{pages}{064001} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.86.064001}.

\bibitem[{\citenamefont{K{\"u}mmel et~al.}(1978)\citenamefont{K{\"u}mmel,
  L{\"u}hrmann, and Zabolitzky}}]{kummel1978}
\bibinfo{author}{\bibfnamefont{H.}~\bibnamefont{K{\"u}mmel}},
  \bibinfo{author}{\bibfnamefont{K.~H.} \bibnamefont{L{\"u}hrmann}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{J.~G.}
  \bibnamefont{Zabolitzky}}, \bibinfo{journal}{Phys. Rep.}
  \textbf{\bibinfo{volume}{36}}, \bibinfo{pages}{1} (\bibinfo{year}{1978}).

\bibitem[{\citenamefont{Day and Zabolitzky}(1981)}]{day1981}
\bibinfo{author}{\bibfnamefont{B.~D.} \bibnamefont{Day}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{G.}~\bibnamefont{Zabolitzky}},
  \bibinfo{journal}{Nucl. Phys. A} \textbf{\bibinfo{volume}{366}},
  \bibinfo{pages}{221} (\bibinfo{year}{1981}).

\bibitem[{\citenamefont{Carlson et~al.}(2003)\citenamefont{Carlson, Morales,
  Pandharipande, and Ravenhall}}]{carlson2003}
\bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Carlson}},
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Morales}},
  \bibinfo{author}{\bibfnamefont{V.~R.} \bibnamefont{Pandharipande}},
  \bibnamefont{and} \bibinfo{author}{\bibfnamefont{D.~G.}
  \bibnamefont{Ravenhall}}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{68}}, \bibinfo{pages}{025802}
  (\bibinfo{year}{2003}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.68.025802}.

\bibitem[{\citenamefont{Gandolfi et~al.}(2009)\citenamefont{Gandolfi,
  Illarionov, Schmidt, Pederiva, and Fantoni}}]{gandolfi2009}
\bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{A.~Y.} \bibnamefont{Illarionov}},
  \bibinfo{author}{\bibfnamefont{K.~E.} \bibnamefont{Schmidt}},
  \bibinfo{author}{\bibfnamefont{F.}~\bibnamefont{Pederiva}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Fantoni}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{79}},
  \bibinfo{pages}{054005} (\bibinfo{year}{2009}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.79.054005}.

\bibitem[{\citenamefont{Gezerlis and Carlson}(2010)}]{gezerlis2010}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Gezerlis}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{J.}~\bibnamefont{Carlson}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{81}},
  \bibinfo{pages}{025803} (\bibinfo{year}{2010}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.81.025803}.

\bibitem[{\citenamefont{Lovato et~al.}(2012)\citenamefont{Lovato, Benhar,
  Fantoni, and Schmidt}}]{lovato2012}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Lovato}},
  \bibinfo{author}{\bibfnamefont{O.}~\bibnamefont{Benhar}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Fantoni}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{K.~E.} \bibnamefont{Schmidt}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{024003} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.85.024003}.

\bibitem[{\citenamefont{Gezerlis et~al.}(2013)\citenamefont{Gezerlis, Tews,
  Epelbaum, Gandolfi, Hebeler, Nogga, and Schwenk}}]{gezerlis2013}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Gezerlis}},
  \bibinfo{author}{\bibfnamefont{I.}~\bibnamefont{Tews}},
  \bibinfo{author}{\bibfnamefont{E.}~\bibnamefont{Epelbaum}},
  \bibinfo{author}{\bibfnamefont{S.}~\bibnamefont{Gandolfi}},
  \bibinfo{author}{\bibfnamefont{K.}~\bibnamefont{Hebeler}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Nogga}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Schwenk}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{111}},
  \bibinfo{pages}{032501} (\bibinfo{year}{2013}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.111.032501}.

\bibitem[{\citenamefont{Dickhoff and Barbieri}(2004)}]{dickhoff2004}
\bibinfo{author}{\bibfnamefont{W.}~\bibnamefont{Dickhoff}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Barbieri}},
  \bibinfo{journal}{Progress in Particle and Nuclear Physics}
  \textbf{\bibinfo{volume}{52}}, \bibinfo{pages}{377 } (\bibinfo{year}{2004}),
  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0146641004000535}.

\bibitem[{\citenamefont{Som\`a and Bo\ifmmode~\dot{z}\else
  \.{z}\fi{}ek}(2008)}]{soma2008}
\bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Som\`a}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{P.}~\bibnamefont{Bo\ifmmode~\dot{z}\else
  \.{z}\fi{}ek}}, \bibinfo{journal}{Phys. Rev. C}
  \textbf{\bibinfo{volume}{78}}, \bibinfo{pages}{054003}
  (\bibinfo{year}{2008}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevC.78.054003}.

\bibitem[{\citenamefont{Rios and Som\`a}(2012)}]{soma2012}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Rios}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{V.}~\bibnamefont{Som\`a}},
  \bibinfo{journal}{Phys. Rev. Lett.} \textbf{\bibinfo{volume}{108}},
  \bibinfo{pages}{012501} (\bibinfo{year}{2012}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevLett.108.012501}.

\bibitem[{\citenamefont{Carbone et~al.}(2013)\citenamefont{Carbone, Polls, and
  Rios}}]{carbone2013}
\bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Carbone}},
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Polls}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{A.}~\bibnamefont{Rios}},
  \bibinfo{journal}{Phys. Rev. C} \textbf{\bibinfo{volume}{88}},
  \bibinfo{pages}{044302} (\bibinfo{year}{2013}).



\bibitem[{\citenamefont{Gros}(1992)}]{gros1992}
\bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Gros}},
  \bibinfo{journal}{Zeitschrift f{\"u}r Physik B Condensed Matter}
  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{359} (\bibinfo{year}{1992}),
  ISSN \bibinfo{issn}{0722-3277},
  \urlprefix\url{http://dx.doi.org/10.1007/BF01323728}.

\bibitem[{\citenamefont{Gros}(1996)}]{gros1996}
\bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Gros}}, \bibinfo{journal}{Phys.
  Rev. B} \textbf{\bibinfo{volume}{53}}, \bibinfo{pages}{6865}
  (\bibinfo{year}{1996}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevB.53.6865}.

\bibitem[{\citenamefont{Lin et~al.}(2001)\citenamefont{Lin, Zong, and
  Ceperley}}]{lin2001}
\bibinfo{author}{\bibfnamefont{C.}~\bibnamefont{Lin}},
  \bibinfo{author}{\bibfnamefont{F.~H.} \bibnamefont{Zong}}, \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{D.~M.} \bibnamefont{Ceperley}},
  \bibinfo{journal}{Phys. Rev. E} \textbf{\bibinfo{volume}{64}},
  \bibinfo{pages}{016702} (\bibinfo{year}{2001}),
  \urlprefix\url{http://link.aps.org/doi/10.1103/PhysRevE.64.016702}.

\bibitem{hagen2014} G.~Hagen {\em et al}, Phys.~Rev.~C {\bf 89}, 01431a9 (2014).

\bibitem{baardsen2014} G.~Baardsen, PhD thesis, University of Oslo (2014) 
\url{https://www.duo.uio.no/handle/10852/41025}. 

\bibitem[{\citenamefont{Shavitt and Bartlett}(2009)}]{shavittbartlett2009}
\bibinfo{author}{\bibfnamefont{I.}~\bibnamefont{Shavitt}} \bibnamefont{and}
  \bibinfo{author}{\bibfnamefont{R.~J.} \bibnamefont{Bartlett}},
  \emph{\bibinfo{title}{Many-body Methods in Chemistry and Physics}}
  (\bibinfo{publisher}{Cambridge University Press}, \bibinfo{year}{2009}).

\bibitem{st83} S.L.\ Shapiro and S.A.\ Teukolsky,
               Black Holes, White Dwarfs and Neutron Stars,
               (Wiley, New York, 1983), Kap 2-4, 5, 8-11, appendiks F

\bibitem{pplp92} M.\ Prakash, M.\ Prakash, J.M.\ Lattimer and C.J.\
Pethick, Astrophys.\ J.\  390 (1992) L77.

\bibitem{pethick92} C.J.\ Pethick, Rev.\ Mod.\ Phys.\ 64 (1992) 1133.

\bibitem{prakash94} M.\ Prakash, Phys.\ Rep.\ 242 (1994) 191.


\bibitem{iwamoto82} N.\ Iwamoto,\ Ann.\ Phys.\  141 (1982) 1.

\bibitem{cs64} H.-Y.\ Chiu and E.E.\ Salpeter, Phys.\ Rev.\ Lett.\
12 (1964) 413.
\bibitem{it72} I.\ Itoh and T.\ Tsuneto, Prog.\ Theor.\ Phys.\ 48 (1972) 149.

\bibitem{fm79} B.L.\ Friman and O.V.\ Maxwell, 
               Astrophys.\ J.\ 232 (1979) 541.


\bibitem{glendenning91} N.K.\ Glendenning, Compact Stars, (Springer, Berlin, 1997).
\bibitem{pbpelk97} M.\ Prakash, I.\ Bombaci, M.\ Prakash, P.J.\ Ellis, J.M.\ 
Lattimer and R.\
Knorren, Phys.\ Rep.\ 280 (1997) 1. 
\bibitem{tsuruta98} S.\ Tsuruta, Phys.\ Rep.\ 292 (1998) 1.

\bibitem{pal97} L.B.\ Leinson and A.\ P\'erez, JHEP 9 (1998) 20;
S.\ Chakrabarty, D.\ Bandyopadahay and S.\ Pal, Phys.\ Rev.\ Lett.\
78 (1997) 2898; ibid.\ 79 (1997) 2176.
\bibitem{HH99} H.\ Heiselberg and M.\ Hjorth-Jensen, 
                      Phys. Rep. 328 (2000) 237.
\bibitem{page94} D.\ Page, Astrophys.\ J.\ 428 (1994) 250.
\bibitem{schaab97} Ch.\ Schaab, D.\ Voskresensky, A.D.\ Sedrakian,
                   F.\ Weber and M.K.\ Weigel, Astron.\ Astrophys.\
                   321 (1997) 591; Ch.\ Schaab, 
                   F.\ Weber, M.K.\ Weigel and N.K.\ Glendenning, 
                   Nucl.\ Phys.\ A 605 (1996) 531.
\bibitem{schaab98} Ch.\ Schaab, F.\ Weber and M.K.\ Weigel, 
                   Astron.\ Astrophys.\ 335 (1998) 596.
\bibitem{malfliet} H.F.\ Boersma and R.~Malfliet, Phys.~Rev.~C 49 (1994) 633.

\bibitem{armen1} Armen Sedrakian and Alex Dieperink, 
Phys.~Lett.~B463 (1999) 145.
\bibitem{armen2}   Armen Sedrakian, preprint nucl-th/0601086, 
http://arxiv.org/abs/nucl-th/0601086.
\end{thebibliography}



\end{document}









