% Chapter Template

\stepcounter{section}

\chapter{Introduction} % Main chapter title

\label{Preface} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\lhead{\emph{Introduction}} % Change X to a consecutive number; this is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
%	Theory
%----------------------------------------------------------------------------------------

In this thesis we will use the coupled cluster method, also called
coupled cluster \emph{theory}, to investigate the homogeneous electron
gas. More specifically, we will investigate how the inclusion of triple
excitations in the coupled cluster equations will affect calculations
of the ground state energy.  Numerous (see for example
Refs.~\cite{Baardsen2014, Shepherd2012, Roggero2013}) calculations of
the system in question have been made using the coupled cluster doubles
(CCD) method, doubles with perturbative triples (CCD(T)), and various
other many-body methods, but to the author's best knowledge this
thesis presents for the first time results containing triple
excitations for the electron gas.

We shall systematically investigate how each diagram present in the
equation contributes to the ground state energy for smaller basis
sets, incrementally leading up the the full inclusion of triple
excitations (CCDT), and we will estimate the energy in the
thermodynamical limit using a CCDT-1 \cite{ShavittBartlett2009}
approach and compare these results with existing results from full
configuration interaction quantum monte carlo (FCIQMC) present in the
literature.  Furthermore, we have also derived a software which allows for the automatical derivation
of various coupled cluster truncations, with the derivation of the corresponding analytical
expressions, their pertinent diagrams and code.

In order to convey these results, this thesis is structured as follows:
In the second chapter of this thesis, we review the
context from which the many-body problem arises, and we motivate the
need for the formalism developed in Chapters \ref{Chapter2}
and \ref{Chapter3}. In Chapter \ref{Chapter4}, we give a broad
overview of the methodology involved, and we give some special
attention to the methods that closely relate to the coupled cluster
method, such as Hartree-Fock theory, configuration interaction theory
and many-body perturbation theory.

In Chapter \ref{Chapter5}, we derive the coupled cluster method, and
we introduce both the pertinent diagrammatic techniques and the computational
implementations that we use to derive the actual equations
for various truncations of the coupled cluster method. Furthermore, we discuss also
our software for deriving automatically various coupled cluster truncations. This software allows for a
consistent check of all derived equations.

The homogeneous electron gas (HEG) is reviewed in
Chapter \ref{Chapter6}, where we derive and present the expressions needed
to evaluate the coupled cluster equations, such as the the plane wave
basis, Fock matrix elements \cite{Thijssen}, the two-body
interaction \cite{ShavittBartlett2009} and the general structure of
the model. 

In Chapter \ref{Chapter7} we present implementational details, and we
describe two conceptually different schemes for solving the coupled
cluster equations for our system. We also discuss topics such as
performance and memory usage.

In Chapter \ref{Chapter8}, we validate the implementations
by comparing with results published in other studies and by comparing
results from the two independent solvers. We then perform a series of
calculations for smaller basis sets, ending up with the full inclusion
of triple amplitudes. Estimates in the thermodynamical limit is
performed by extrapolation from a data set obtained by running our
CCDT-1 code on the Abel cluster (see Ref.~\cite{abel}).

We discuss these findings in light of calculations performed by
others, and finally, in the last chapter,  we present  our concluding remarks, recommendations
and perspectives for future work.












