\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
The completion of this thesis marks the end of a story that begun almost twenty years ago.
As a small boy, barely six years old and just started at elementary school, I pondered everything that crossed my mind.
Standing outside our cottage in the winter watching the stars, and occasionally also the northern light, I was stunned by the size and the beauty of the universe we live in.
Later, as I watched wood burn in the fireplace, I enjoyed the stories told me about how the sun's warmth was stored inside the logs, to be released in winter by lighting a match.

 
The decision of going in the direction of natural sciences came early, initially made by myself, but it would not have been realized if I had not gotten the help by people I met on my way.
Whenever I had questions my dad helped feed my curiosity, and whenever my curiosity let me down my mother gave me the care and comfort I needed.

 
Throughout years of education I have had both positive and negative experiences.
The best experience was meeting Morten Hjorth-Jensen as the teacher in computational physics.
I know no other professor with the same enthusiasm and excellence, both academic and pedagogical, and there is no coincidence why I returned half a year later to discuss topics for a Master's Thesis with him.
Having an office available during the Master I spent more time working at the university, and I appreciate the collegues who sat close by.
I would like to mention Frank Olsen (we spent some time discussing different topics regarding our similar theses), Jørgen Høgberget (another intelligent, but out-of-control, coffee addict), and Sarah Reimann (learning me German cannot have been easy, good thing we focused mainly on one word).

 
The rest of my family also deserves a few words here. 
My sisters, who has the courage to be there for me anytime, even in times when I should have been there for them instead, and my brother, who helped proof reading this thesis.
At last, but not at all the least, I appreciate how my girlfriend, soon my wife, is still there, despite all the days and nights I have spent in front of my computer.
It is no exaggeration to say that I am proud of my family.

 
To all of you, and others who has helped me during all this time, I owe you my thanks.
Hopefully have I lived up to your expectations as well.
I am looking forward to spend more time with you this summer, before I may resume my story or start a new one.




\begin{flushright}
Christoffer Hirth
\end{flushright}

\begin{tabular}{c}
Oslo,\\
June, 2012
\end{tabular}



