\documentclass[12pt]{report}
\begin{document}
\thispagestyle{empty}
\center{Enclosure (a): CURRICULUM VITAE}
\vspace{10mm}
\flushleft

{\it Personal data:}

\begin{description}
\item{Name : Elise Bergli}
\item{Born : 6. July 1977}
\item{Postal address : Schouterrassen 6,

\hspace{15mm}0573 OSLO}
\item{Phone : 92 05 19 98}
\item{Civil state : engaged to be married summer 2002}


\end{description}

\vspace{10mm}

{\it Education :}

\begin{description}
\item{1984-1990 Primary School in Kongsberg}
\item{1993-1996 Kongsberg Videregående Skole (High School in Kongsberg). 

Specializing in Mathematics, Physics, Chemistry and French}

\item{1996-1999 University of Oslo, Faculty of Mathematics and Sciences and Faculty of Arts. 

Interdisciplinary Bachelor-degree in Physics and Logic, Language and Information (SLI).}

\item{1999-2001 University of Oslo, Faculty of Mathematics and Sciences. 

Master Degree in Physics, at the Solid State Department. Thesis titled ``A Simple Model for the Physics of Surface Contact and Adhesion''. Advisors Jens Feder and Torstein J{\o}ssang.}

\item{2001-\hspace{10mm}University of Oslo, Faculty of Arts. 

Currently working on a Master Degree in Logic, Language and Information (SLI), on formal languages.}

\end{description}
\vspace{10mm}

{\it Previous Job Engagements :}

\begin{description}

\item{Summer 1997 Assistant nurse at Skavangertun Eldre- og Sykehjem}
\item{Spring 1998 Assistant teacher in SLI1 (Logic and Computation I) at the University of Oslo}
\item{Spring 1999 Assistant teacher in SLI110 (Logic and Computation I) at the University of Oslo}
\item{Summer 1999 Lecturer in Physics and Mathematics, High School level (2FY, 2MX, 3MX) at Buskerud University College}
\item{Spring 2000 Assistant teacher in SLI110 (Logic and Computation I) at the University of Oslo}
\item{Summer 2000 Lecturer in Physics and Mathematics, High School level (2FY, 2MX, 3MX) at Buskerud University College}
\item{Autumn 2000 Assistant teacher in SLI260 (Computational Linguistics I) at the University of Oslo}
\item{Summer 2001 Lecturer in Physics and Mathematics, High School level (2FY, 2MX, 3MX) at Buskerud University College}
\item{Autumn 2001 Assistant teacher in SLI110 (Logic and Computation I) at the University of Oslo}
\end{description}

\end{document}
