% Introduction
Graduate school has been an amazing journey and, luckily, a journey that I have not had to take alone. All the people listed below have contributed in some way and helped make this dissertation a reality. It has been a dream of mine since I was in high school to get a Ph.D.; there is no way I could have done it without these fantastic people.

% Morten
There is no other way to start this acknowledgments section than by extending a huge thank you to my advisor, Morten Hjorth-Jensen.  Morten is the best teacher and mentor I could have asked for. He has influenced my research, teaching, and outlook on the world more than I could have imagined. Morten, while I may no longer be your student, I will always look to you as an example of what a professor and a research advisor should be. 

% Committee and other MSU
I want to thank my committee members Scott Bogner, Johannes Pollanen, Danny Caballero, and Sean Liddick for their unending patience and assistance throughout this process. I also must thank Heiko Hergert for introducing me to many-body theory during my first summer at MSU and aiding me throughout this journey. I have also been lucky to share this journey as a graduate student with my group members, Yani Udiani, Jane Kim, Danny Jammooa, and Kang Yu. Being able to experience all of the aspects of being a nuclear theory graduate student with you has been a pleasure.

% Justin and Gustav
I would also like to acknowledge my collaborators at Oak Ridge National Laboratory, Gustav Jansen and Justin Leitz, without whom the coupled cluster results presented in this thesis would not be possible. Not only did you make significant contributions to my research, but I also look up to you as scientists and mentors. Both of you made my time living in Oak Ridge much more manageable and enjoyable.

% Elizabeth and Kim
Next, I want to acknowledge Kim Crosslan and Elizabeth Delinski for all of their assistance over the past five years.  There is no way I could have navigated all of the different aspects of graduate school without your help.

% Family
Graduate school would have been much harder without the encouragement and support of my family. I have to thank my parents, Jimmy and Linda, for their support throughout this process. My sister, Bre, has been an unending source of encouragement when things got hard, and my brother-in-law, Graham, has been the best editor I could ask for. I must also acknowledge the family I gained through marriage: Chris, Mary Jane, Aaron, Nana, Papa, Nannie, and Pawpaw. There is no way I could have done all I have over the past five years if you were not there to support me.

% Friends and Kelvin 
I want to thank the friends that have been with me since college (Courtney, Connor, Mikhayla, Will) and the friends I have made in graduate school (Eric, Nick, and Ana) for always encouraging me and believing in me even when I did not believe in myself. Graduate school would have been a much more lonely place without all of the company and good times you provide. 

I also have to acknowledge to contributions of my dog, Lord Kelvin, to my mental health and sanity throughout all of the virtual classes and work-from-home days I have endured as a graduate student.

% Adam
Finally, I must acknowledge the person this thesis is dedicated to: my husband, Adam. There is no way I can ever pay you back for all of the assistance you have provided and the sacrifices you have made to help me get this degree, but I dedicate this thesis to you as an acknowledgment of everything you have done.
