\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
The completion of this thesis marks the end of a story that begun more than 15 years ago.
As a small boy, barely started at elementary school, I pondered everything that crossed my mind.
Standing outside our cottage in the winter watching the stars, and occasionally also the northern light, I was stunned by the size and the beauty of the universe we live in.
Later, as I watched wood burn in the fireplace, I enjoyed the stories told me about how the sun's warmth was stored inside the logs, to be released in winter by lighting a match.

 
The decision of going in the direction of natural sciences came early, initially made by myself, but it would not have been realized if I had not gotten help from people I met on my way.
Whenever I had questions my dad helped feed my curiosity, and whenever my curiosity let me down my mother gave me the care and comfort I needed.

 
Throughout years of education I have had both positive and negative experiences.
The best experience was meeting Morten Hjorth-Jensen as the teacher in computational physics.
I know no other professor with the same enthusiasm and excellence, both academic and pedagogical, and there is no coincidence why I returned half a year later to discuss topics for a Master's Thesis with him.
Having an office available during the Master I spent more time at the university, and I appreciate the collegues who sat close by.
I would like to mention Frank Olsen (we spent some time discussing different topics regarding our similar theses), Jørgen Høgberget (your fresh grind coffee lights up the day), and Sarah Reimann (learning me German cannot have been easy, good thing we focused mainly on one word).

 
The rest of my family also deserves a few words here. 
My sisters, who have the courage to always be there for others, even in times when others should have been there for them instead, and my brother, who is always positive, even when helping me proof read this thesis.
At last, but not at all the least, I appreciate how my girlfriend, soon my wife, is still there, despite all the days and nights I have spent in front of my computer.
It is no exaggeration to say that I am proud of my family.

 
To all of you, and others who have helped me during all this time, I owe you my thanks.
Hopefully have I lived up to your expectations, and I look forward to spending more time with you this summer, before I may resume my story or start a new one.




\begin{flushright}
Christoffer Hirth
\end{flushright}

\begin{tabular}{c}
Oslo,\\
June, 2012
\end{tabular}



